---
title: "My Dataset"
output: "github_document"
date: "2024-09-26"
---
```{r}
# Set a CRAN mirror
options(repos = c(CRAN = "https://cran.rstudio.com/"))

```

```{r}
#install.packages("dplyr")
library(car)
library(dplyr) #recode variables
library(tidyr)
library(ggplot2)
library(psych)
library(bruceR)
library(haven) #load CSV
library(Rmisc)
library(tidyverse)
library(bruceR)
#library(haven)
#install.packages("nortest")

#ibrary(nortest)
#install.packages("car")
```
Reading dataset 
```{r}

# Read the TSV file
data <- read.table("C:/Users/sidne/OneDrive/Desktop/PSY329/38964-0001-Data.tsv", sep = "\t", header = TRUE, fill = TRUE, quote = "")

# Write the data to a CSV file
write.csv(data, "Excel-Data.csv", row.names = FALSE)

#open file
data <- read.csv("C:/Users/sidne/OneDrive/Desktop/PSY329/Excel-Data.csv")

```

Cleaning data set and recoding variables
```{r}
clean_data <- data %>%
  select(SEX,HHINC,HAPPY,SAT1,SAT2,SAT3,SAT4,SAT5,LONELY_A,LONELY_B,LONELY_C) 

clean_data$SEX<-recode(data$SEX, '1' = 'Male', '2' = 'Female', '3'='Other')



clean_data$HOUSE_INCOME<-recode(data$HHINC, '1' = 'Low', '2' = 'Low', '3'='Low','4'='Low','5'='Low','6'='Middle','7'='Middle','8'='Middle','9'='Middle','10'='Middle','11'='High','12'='High')

clean_data$HAPPY_meaning<-recode(data$HAPPY, '1' = 'Very happy', '2' = 'Rather happy', '3'='Not very happy','4'='Not at all happy')

clean_data <- clean_data %>%
   mutate(Lonely=rowMeans(cbind(LONELY_A,LONELY_B,LONELY_C))) #1 hardly ever feel lonely to 5 often feel lonely 

clean_data <- clean_data %>%
  mutate(Lonely = ifelse(Lonely > 3, NA, Lonely)) #removing participants that responded don't know or refused the question

clean_data <- clean_data %>%
   mutate(Life_satisfaction=rowMeans(cbind(SAT1,SAT2,SAT3,SAT4,SAT5))) #1 strongly disagree to 5 strongly agree higher score = higher life satisfaction

clean_data <- clean_data %>%
  mutate(Life_satisfaction = ifelse(Life_satisfaction > 5, NA, Life_satisfaction))

clean_data$HAPPY_REV <- 5 - clean_data$HAPPY #reverse code happy so higher number = happier

clean_data <- clean_data %>%
  select(SEX,HOUSE_INCOME,HHINC,HAPPY_meaning, HAPPY_REV,Life_satisfaction, SAT1,SAT2,SAT3,SAT4,SAT5,Lonely, LONELY_A,LONELY_B,LONELY_C)

clean_data<- drop_na(clean_data)
clean_data <- clean_data %>% filter(SEX != "Other")

write.csv(clean_data, "C:/Users/sidne/OneDrive/Desktop/PSY329/clean_data.csv", row.names = FALSE)

```

Normality of Lonely
```{r}

ggplot(clean_data, aes(x = Lonely)) + geom_histogram(binwidth = 0.5) + facet_wrap(~SEX)+theme_classic() #between SEX

ggplot(clean_data, aes(x = Lonely)) + geom_histogram(binwidth = 0.5) + facet_wrap(~HOUSE_INCOME)+theme_classic() #between INCOME

#clean_data %>%
 # group_by(SEX) %>% 
  #summarize(W = shapiro.test(Lonely)$statistic, p_value = shapiro.test(Lonely)$p.value)

#clean_data %>%
 # group_by(HOUSE_INCOME) %>% 
  #summarize(W = shapiro.test(Lonely)$statistic, p_value = shapiro.test(Lonely)$p.value)

describeBy(Lonely ~ SEX, data= clean_data)
describeBy(Lonely ~ HOUSE_INCOME, data= clean_data)
```

Normality of Life satisfaction
```{r}
ggplot(clean_data, aes(x = Life_satisfaction)) + geom_histogram(binwidth = 0.5) + facet_wrap(~SEX)+theme_classic()

ggplot(clean_data, aes(x = Life_satisfaction)) + geom_histogram(binwidth = 0.5) + facet_wrap(~HOUSE_INCOME)+theme_classic()

#clean_data %>%
 # group_by(SEX) %>% 
  #summarize(W = shapiro.test(Life_satisfaction)$statistic, p_value = shapiro.test(Life_satisfaction)$p.value)

#clean_data %>%
 # group_by(HOUSE_INCOME) %>% 
  #summarize(W = shapiro.test(Life_satisfaction)$statistic, p_value = shapiro.test(Life_satisfaction)$p.value)

describeBy(Life_satisfaction ~ SEX, data= clean_data)
describeBy(Life_satisfaction ~ HOUSE_INCOME, data= clean_data)

```

Normality of Happiness
```{r}
ggplot(clean_data, aes(x = HAPPY_REV)) + geom_histogram(binwidth = 1) + facet_wrap(~SEX)+theme_classic()

ggplot(clean_data, aes(x = HAPPY_REV)) + geom_histogram(binwidth = 1) + facet_wrap(~HOUSE_INCOME)+theme_classic()

#clean_data %>%
 # group_by(SEX) %>% 
  #summarize(W = shapiro.test(HAPPY_REV)$statistic, p_value = shapiro.test(HAPPY_REV)$p.value)

#clean_data %>%
 # group_by(HOUSE_INCOME) %>% 
  #summarize(W = shapiro.test(HAPPY_REV)$statistic, p_value = shapiro.test(HAPPY_REV)$p.value)

describeBy(HAPPY_REV ~ SEX, data= clean_data)
describeBy(HAPPY_REV ~ HOUSE_INCOME, data= clean_data)

```
Variance of Lonely
```{r}
clean_data %>%
  group_by(SEX) %>%
  summarize(variacne = var(Lonely))

leveneTest(Lonely~SEX, clean_data)

clean_data %>%
  group_by(HOUSE_INCOME) %>%
  summarize(variacne = var(Lonely))

leveneTest(Lonely~HOUSE_INCOME, clean_data)
```
Variance of Life satisfaction
```{r}
clean_data %>%
  group_by(SEX) %>%
  summarize(variacne = var(Life_satisfaction))

leveneTest(Life_satisfaction~SEX, clean_data)

clean_data %>%
  group_by(HOUSE_INCOME) %>%
  summarize(variacne = var(Life_satisfaction))

leveneTest(Life_satisfaction~HOUSE_INCOME, clean_data)

```
Variance of Happiness
```{r}
clean_data %>%
  group_by(SEX) %>%
  summarize(variacne = var(HAPPY_REV))

leveneTest(HAPPY_REV~SEX, clean_data)

clean_data %>%
  group_by(HOUSE_INCOME) %>%
  summarize(variacne = var(Life_satisfaction))

leveneTest(HAPPY_REV~HOUSE_INCOME, clean_data)
```
Life satisfaction ANOVA
```{r}
mod<-MANOVA(clean_data, dv = "Life_satisfaction", between = c("SEX", "HOUSE_INCOME")) 

EMMEANS(mod, effect = "SEX", by = "HOUSE_INCOME", p.adjust = "none")

EMMEANS(mod, effect = "HOUSE_INCOME", by = "SEX", p.adjust = "none")
```
Loneliness
```{r}
mod<-MANOVA(clean_data, dv = "Lonely", between = c("SEX", "HOUSE_INCOME")) 

EMMEANS(mod, effect = "SEX", by = "HOUSE_INCOME", p.adjust = "none")

EMMEANS(mod, effect = "HOUSE_INCOME", by = "SEX", p.adjust = "none")
```
Graphs
```{r}

plot<-summarySE(clean_data, measurevar="Life_satisfaction", groupvars=c("SEX", "HOUSE_INCOME")) #to get mean scores before putting into 
ggplot(plot, aes(x = SEX, y = Life_satisfaction, fill = SEX)) +
  geom_col() + facet_wrap(~ HOUSE_INCOME) + theme_bruce()

plot2<-summarySE(clean_data, measurevar="Lonely", groupvars=c("SEX", "HOUSE_INCOME"))
ggplot(plot2, aes(x = SEX, y = Lonely, fill = SEX)) +
  geom_col() + facet_wrap(~ HOUSE_INCOME) + theme_bruce()

#plot3<-summarySE(clean_data, measurevar="Lonely", groupvars=c("SEX", "HOUSE_INCOME"))

ggplot(plot, aes(x = HOUSE_INCOME, y = Life_satisfaction, fill = HOUSE_INCOME)) +
  geom_col() + facet_wrap(~ SEX) + theme_bruce()

ggplot(plot2, aes(x = HOUSE_INCOME, y = Lonely, fill = HOUSE_INCOME)) +
  geom_col() + facet_wrap(~ SEX) + theme_bruce()

```
```{r}
clean_data$HHINC_n<-as.numeric(clean_data$HHINC)

PROCESS(clean_data, y = "Life_satisfaction", x = "HHINC_n", mods = c("SEX"))
ggplot(clean_data, aes(x = HHINC_n, y = Life_satisfaction)) + geom_point() + geom_smooth() + theme_bruce() +facet_wrap("SEX")

PROCESS(clean_data, y = "Lonely", x = "HHINC_n", mods = c("SEX"))
ggplot(clean_data, aes(x = HHINC_n, y = Lonely)) + geom_point() + geom_smooth() + theme_bruce() +facet_wrap("SEX")

```
Reliability and Factor analysis for  Satisfaction with Life
```{r}
Alpha(clean_data, "SAT", 1:5) 

EFA(clean_data, "SAT", 1:5, method = "pa", plot.scree = TRUE, nfactors = c("parallel"))
```

Reliability and Factor analysis for Lonliness
```{r}

Alpha(clean_data,"LONELY_", c("A","B","C"))

EFA(clean_data, "LONELY_", c("A","B","C"), method = "pa", plot.scree = TRUE, nfactors = c("parallel"))
```




